---
title: "mahi"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mahi}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(multimediate)
```


Simulation of high dimensional data

```{r}
set.seed(42)
nbrsamples <- 1 # number of datasets to simulate
n <- 100 # number of observationsn , n = 300
p <- 500#0 # number of mediators, p = 500, 1000 and 5000
alpha <- rep(0,p) # exposure --> mediators
beta <- rep(0,p) # mediators --> outcome
gamma <- 0.5 # exposure --> outcome
c <- 0.2 # intercept term in the outcome model

# true mediators

# 10 easy to predict
alpha[1:10] <- c(-0.95, 0.90, 0.95, 0.95, -1.0, 0.95, -0.95, 0.85, -0.95, -1.0)*10
beta[1:10] <- c(1.85, -1.95, 1.90, 1.85, -1.90, 1.85, 1.95, -1.90, 1.0, 1.85)*10

# 10 middle easy to predict
alpha[11:20] <- c(-0.55, 0.65, -0.70, 0.50, -0.75, 0.60, -0.65, 0.55, 0.60, 0.70)*5
beta[11:20] <- c(0.65, -0.50, 0.75, -0.60, 0.70, -0.55, 0.75, -0.65, 0.65, -0.55)*5

# 10 hard to predict
alpha[21:30] <- c(-0.35, 0.25, -0.15, -0.10, 0.20, -0.35, 0.05, -0.45, 0.15, -0.10)*2.5
beta[21:30] <- c(0.15, -0.10, 0.25, -0.45, 0.30, 0.10, -0.20, 0.15, -0.25, 0.30)*2.5

# not true mediators
alpha[31:40] <- 0.25*10 # betas are zero
beta[41:50] <- 0.45*10 # alphas are zero

simdatasets <- list()

# In the following : X = exposure, M = mediators, Y = outcome, C = confounder
for (s in 1:nbrsamples) {

  ck <- t(runif(p, -2, 2)) # intercept term of the mediator models
  M <- matrix(0, n, p) # matrix that contains the mediators
  X <- ifelse(runif(n, 0, 1) < 0.5, 0, 1) # binary exposure vector
  #C <- rnorm(n, 0, 0) # confounder effect (only one for now)
  #ksi1 <- rnorm(p, 0, 0) # regression coefficient of confounder effect --> M, 
  # different coeff in front of each mediator, but not different from one individual to the other

  for (i in 1 : n) {
    e <- rnorm(p, 0, 1.5) # error term of the mediator models
    M[i,] <- ck + X[i] * alpha + e#+ C[i]*ksi1 + e  
    # Each mediator model contains one confounder effect,
    #same value for all mediators p, but different value from an individual to the other
  }
  colnames(M) <- paste0("M", 1:ncol(M))

  XM <- cbind(X, M)  #  [X M]
  epsilon <- rnorm(n, 0, 1.5) # error term of the outcome model
  ksi2 <- 0.1 # regression coefficient of confounder effect --> Y
  B <- c(gamma, beta)  # (p+1) times 1
  Y <- c + XM %*% t(t(B)) + t(t(epsilon))#+ C*ksi2 + t(t(epsilon)) #  the response vector
  #Y <- matrix(rbinom(n, 1, 1/(1 + exp(-Y))), nrow = n) # binary outcome

simdatasets[[s]] <- cbind(X, M, Y)#, C)
}
   
for (s in 1:nbrsamples) {
  simdatasets[[s]] <- as.data.frame(simdatasets[[s]])
  colnames(simdatasets[[s]])[1] <- "X" # renaming the exposure as X
  colnames(simdatasets[[s]])[p+2] <- "Y" # renaming the outcome as Y
}
```


Before using the function mahi we have to choose a value for lambda using lambdaChoice function. We have to set a max for lambda with lambdamax. The returned value is less than lambdamax and allows you to select a mediator number between selectedMin and selectedMax. The value of selectedMax must be less than or equal to the number of individuals n. If the value of lambdamax is not hifh enough you will have this message. 

```{r}
datahigh=simdatasets[[1]]
n=dim(datahigh)[1] #sample number
P=1 #number of treatment
treatment=matrix(datahigh$X,n,P)
mediators=as.matrix(datahigh[,paste0("M", 1:p)])
outcome=datahigh$Y


  lambda <- lambdaChoice(datahigh,name.exposure="X",name.outcome="Y",name.mediators=paste0("M", 1:p),lambdamax=10,selectedMin=n/2,selectedMax=n,L0=.1,eta=2,tau=1/P,epsilon=.001)

```

You have to choose a greater value, for example here 300.
```{r}
lambda <- lambdaChoice(datahigh,name.exposure="X",name.outcome="Y",name.mediators=paste0("M", 1:p),lambdamax=500,selectedMin=n/2,selectedMax=n,L0=.1,eta=2,tau=1/P,epsilon=.001)
lambda
```

Now you can use the mahi function for the selection
```{r,include=FALSE}
  mahifit <- mahi(data=datahigh,name.exposure="X",name.outcome="Y",name.mediators=paste0("M", 1:p),
                   lambda=lambda,Nboot=30,L0=.1,eta=2,tau=1/P,epsilon=.001,
                   Kmax=NULL,bin=FALSE,
                   step2=TRUE,p.adjust.method="hochberg",pvalseuil=0.05)
```

```{r,eval=FALSE}
  mahifit <- mahi(data=datahigh,name.exposure="X",name.outcome="Y",name.mediators=paste0("M", 1:p),
                   lambda=lambda,Nboot=30,L0=.1,eta=2,tau=1/P,epsilon=.001,
                   Kmax=NULL,bin=FALSE,
                   step2=TRUE,p.adjust.method="hochberg",pvalseuil=0.05)
```


```{r}
summary(mahifit)
```

The function mahi returns a list containing several results such as the mediators selected in the first step. 


To see the mediators selected after the step 1, use :
```{r}
  mahifit$Kmaxrank
```
Remember that the first 30 candidates are the real mediators. The first 10 (1 to 10) have a strong mediated effect, the next 10 (11 to 20) have a medium mediated effect, and the other 10 (21 to 30) have a very weak effect.


To see the mediators count for each mediator after the stability selection (step 1), use :
```{r}
  mahifit$bootcount
```

To see the ranking after the stability selection (step 1), use :
```{r}
mahifit$ranking
```


To see the results of the multiple analysis with mediators selected at the step 1, use :
```{r}
  summary(mahifit$multimed)
```

You can see the p-value in the previous table or use :
```{r}
  mahifit$pvals
```


The corrected p-value for the multiple test obtained with method given can be disp with :
```{r}
  mahifit$pvalscorr
```
It is then possible to make another correction for a multiple test with the mediators selected in step 1.

To see the mediators selected after step 2 you can use these lines of codes.
```{r}
mahifit$selection

mahiselection <- as.vector(which(mahifit$selection != FALSE))
mahiselection
```
